---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# symbolic

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/symbolic)](https://CRAN.R-project.org/package=symbolic)
<!-- badges: end -->

`symbolic` is an R package designed to provide symbolic algebra capabilities natively in R using R functions.


It is highly experimental and dependent on external contributors to be a fully fledged Computer Algebra System (CAS). It will be developed to the extent needed for my upstream packages, specifically distributional which requires many symbolic algebra tools not yet readily available in R. Currently powered by [yacas](http://www.yacas.org/) via [Ryacas](https://cran.r-project.org/package=Ryacas).


## Installation

You can install the development version of symbolic like so:

``` r
remotes::install_github("mitchelloharawild/symbolic")
```

## Usage

Prepare a function for symbolic transformations with `as_symbolic()`.

```{r}
library(symbolic)
.f <- function(x, y) log(2*x + 1 + y)
.f
.f_sym <- as_symbolic(.f)
.f_sym
```

Internally, we use yacas for all the hard work.

```{r}
.f_yac <- symbolic:::as_yacas(.f)
.f_yac
symbolic:::r_from_yacas(.f_yac)
```

Typical base functions for math have `symbolic` methods which return functions.

```{r}
# log(2 * x + 1 + y) == y # for x
solve(.f_sym, 0, "x")
deriv(.f_sym, "x")(2, 1)
integrate(.f_sym, "x")

.f_sym |> 
  deriv("x") |> 
  solve(y, "x")
```

